

<!----------------------------------------------------------------------------->

---
class: code-marg-5

## Collections and collection members

<p style="position: absolute; bottom:3%; right:10%; margin:0; font-size:small; font-style:italic;">
<a href="https://doi.org/10.1093/nar/gkx1153" target="_blank">
ClinVar: <strong>Landrum et al. (2018)</strong>
</a>
 |
<a href="https://patzaw.github.io/TKCat/TKCat-User-guide.html#25_Documenting_collection_members" target="_blank">
<strong>TKCat user guide</strong> &gt; Collection members
</a>
 | 
<a href="https://patzaw.github.io/TKCat/TKCat-User-guide.html#collections" target="_blank">
<strong>TKCat user guide</strong> &gt; Collections
</a>
</p>

.pull-left[

```{r, eval=FALSE}
clinvar <- read_fileMDB(system.file(
   "examples/ClinVar", package="TKCat"
))
```

```{r, echo=FALSE, results='hide'}
clinvar <- read_fileMDB(system.file("examples/ClinVar", package="TKCat"))
```

```{r, echo=FALSE, results='asis'}
pp <- data_model(clinvar) %>%
   plot() %>%
   visOptions(width="100%", height="260px")
pp$sizingPolicy <- htmlwidgets::sizingPolicy(
   browser.defaultWidth = "500px",
   browser.defaultHeight = "260px",
   browser.padding=0,
   browser.fill = FALSE
)
cvif <- hwIframe(
   p=pp,
   style="height:270px; width:100%; border-style:none; background-color:transparent;"
)
cat(cvif)
```

]

.pull-right[

```{r, echo=FALSE, results='asis'}
pp <- data_model(hpo) %>%
   plot() %>%
   visOptions(width="100%", height="100px")
pp$sizingPolicy <- htmlwidgets::sizingPolicy(
   browser.defaultWidth = "500px",
   browser.defaultHeight = "100px",
   browser.padding=0,
   browser.fill = FALSE
)
hpif <- hwIframe(
   p=pp,
   style="height:120px; width:100%; border-style:none; background-color:transparent;"
)
cat(hpif)
```


]

???

```{r, echo=FALSE, results='asis'}
cat(notes$collections)
```

<!----------------------------------------------------------------------------->

---
class: code-marg-5
count: false

## Collections and collection members

<p style="position: absolute; bottom:3%; right:10%; margin:0; font-size:small; font-style:italic;">
<a href="https://doi.org/10.1093/nar/gkx1153" target="_blank">
ClinVar: <strong>Landrum et al. (2018)</strong>
</a>
 |
<a href="https://patzaw.github.io/TKCat/TKCat-User-guide.html#25_Documenting_collection_members" target="_blank">
<strong>TKCat user guide</strong> &gt; Collection members
</a>
 | 
<a href="https://patzaw.github.io/TKCat/TKCat-User-guide.html#collections" target="_blank">
<strong>TKCat user guide</strong> &gt; Collections
</a>
</p>

.pull-left[

```{r, eval=FALSE}
clinvar <- read_fileMDB(system.file(
   "examples/ClinVar", package="TKCat"
))
```

```{r, echo=FALSE, results='hide'}
clinvar <- read_fileMDB(system.file("examples/ClinVar", package="TKCat"))
```

```{r, echo=FALSE, results='asis'}
cat(cvif)
```

]

.pull-right[

```{r, echo=FALSE, results='asis'}
cat(hpif)
```

```{r}
hpo <- hpo %>% add_collection_member(
   collection="Condition",
   table="HPO_diseases",
   condition=list(
      value="Disease", static=TRUE
   ),
   source=list(
      value="db", static=FALSE
   ),
   identifier=list(
      value="id", static=FALSE
   )
)
```


]

???

```{r, echo=FALSE, results='asis'}
cat(notes$collections)
```

<!----------------------------------------------------------------------------->

---
class: code-marg-5
count: false

## Collections and collection members

<p style="position: absolute; bottom:3%; right:10%; margin:0; font-size:small; font-style:italic;">
<a href="https://doi.org/10.1093/nar/gkx1153" target="_blank">
ClinVar: <strong>Landrum et al. (2018)</strong>
</a>
 |
<a href="https://patzaw.github.io/TKCat/TKCat-User-guide.html#25_Documenting_collection_members" target="_blank">
<strong>TKCat user guide</strong> &gt; Collection members
</a>
 | 
<a href="https://patzaw.github.io/TKCat/TKCat-User-guide.html#collections" target="_blank">
<strong>TKCat user guide</strong> &gt; Collections
</a>
</p>

.pull-left[

```{r, eval=FALSE}
clinvar <- read_fileMDB(system.file(
   "examples/ClinVar", package="TKCat"
))
```

```{r, echo=FALSE, results='hide'}
clinvar <- read_fileMDB(system.file("examples/ClinVar", package="TKCat"))
```

```{r, echo=FALSE, results='asis'}
cat(cvif)
```

```{r, eval=FALSE}
collection_members(clinvar)
```

]

.pull-right[

```{r, echo=FALSE, results='asis'}
cat(hpif)
```

```{r, eval=FALSE}
hpo <- hpo %>% add_collection_member(
   collection="Condition",
   table="HPO_diseases",
   condition=list(
      value="Disease", static=TRUE
   ),
   source=list(
      value="db", static=FALSE
   ),
   identifier=list(
      value="id", static=FALSE
   )
)
```

]

???

```{r, echo=FALSE, results='asis'}
cat(notes$collections)
```



<!----------------------------------------------------------------------------->

---
class: code-marg-5

## Merging with collections

<p style="position: absolute; bottom:3%; right:10%; margin:0; font-size:small; font-style:italic;">
<a href="https://f1000research.com/articles/7-195" target="_blank">
BED: a Biological Entity Dictionary
</a>
 | 
<a href="https://f1000research.com/articles/9-942" target="_blank">
DODO: Dictionary Of Disease Ontologies
 | 
</a>
<a href="https://patzaw.github.io/TKCat/TKCat-User-guide.html#36_Merging_MDBs_with_collections" target="_blank">
<strong>TKCat user guide</strong> &gt; Merging with collections
</a>
</p>

<div class="l64 code-marg-0">

.pull-left[

```{r}
shared_collections <- get_shared_collections(
   hpo, clinvar
)
shared_collections
```

]

</div>

???

```{r, echo=FALSE, results='asis'}
cat(notes$merging)
```

<!----------------------------------------------------------------------------->

---
class: code-marg-5
count: false

## Merging with collections

<p style="position: absolute; bottom:3%; right:10%; margin:0; font-size:small; font-style:italic;">
<a href="https://f1000research.com/articles/7-195" target="_blank">
BED: a Biological Entity Dictionary
</a>
 | 
<a href="https://f1000research.com/articles/9-942" target="_blank">
DODO: Dictionary Of Disease Ontologies
 | 
</a>
<a href="https://patzaw.github.io/TKCat/TKCat-User-guide.html#36_Merging_MDBs_with_collections" target="_blank">
<strong>TKCat user guide</strong> &gt; Merging with collections
</a>
</p>

<div class="l64 code-marg-0">

.pull-left[

```{r}
shared_collections <- get_shared_collections(
   hpo, clinvar
)
shared_collections
```

```{r, eval=FALSE}
hpo_clinvar <- merge(
   hpo, clinvar,
   by=shared_collections[1,]
)
hpo_clinvar %>% data_model() %>% plot()
```

]

.pull-right[

```{r dodoMerge, echo=FALSE, cache=TRUE}
hpo_clinvar <- merge(
   hpo, clinvar,
   by=shared_collections[1,],
   dmAutoLayout=FALSE
)
```


```{r, echo=FALSE, results='asis'}
tp <- hpo_clinvar %>% data_model()
tp$Condition_1_HPO_diseases_1_ClinVar_traitCref$display$x <-
   tp$ClinVar_traitCref$display$x
tp <- unclass(tp)
for(
   n in
   c(
      "HPO_diseases", "HPO_diseaseHP", "HPO_hp",
      "Condition_1_HPO_diseases_1_ClinVar_traitCref"
   )
){
   tp[[n]]$display$y <- -250
}
tp <- RelDataModel(tp)
pp <- tp %>% plot() %>%
   visOptions(width="100%", height="400px")
pp$sizingPolicy <- htmlwidgets::sizingPolicy(
   browser.defaultWidth = "400px",
   browser.defaultHeight = "400px",
   browser.padding=0,
   browser.fill = FALSE
)
cat(hwIframe(
   p=pp,
   style="height:410px; width:100%; border-style:none; background-color:transparent;"
))
```

]

</div>

???

```{r, echo=FALSE, results='asis'}
cat(notes$merging)
```