---
title: "chTKCat user guide"
author: "Patrice Godard"
date:  "`r format(Sys.time(), '%B %d, %Y')`"
output:
   rmarkdown::html_vignette:
     number_sections: yes
     self_contained: yes
     toc: yes
     fig_width: 7
     fig_height: 5
vignette: >
   %\VignetteIndexEntry{chTKCat user guide}
   %\VignetteEncoding{UTF-8}
   %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, message=FALSE, echo=FALSE, include=FALSE, cache=FALSE}
library(knitr)
opts_chunk$set(
   include=TRUE,
   echo=TRUE,
   message=TRUE,
   warning=TRUE,
   cache=FALSE,
   cache.lazy=FALSE
)
## The following line is to avoid building errors on CRAN
knitr::opts_chunk$set(eval=Sys.getenv("USER") %in% c("pgodard"))
library(TKCat)
```

```{r child = 'supp/urls.Rmd'}
```


<!----------------------------------------------------------------------------->
<!----------------------------------------------------------------------------->
# Introduction

```{r child = 'supp/genIntro.Rmd'}
```

This vignette describes how TKCat can be used with
a [ClickHouse][clickhouse] database.
Users should also refer
to the [general TKCat user guide][uguide].
Finally, a [specific vignette][opman] describes
how to create and administrate a TKCat ClickHouse instance.

<!----------------------------------------------------------------------------->
<!----------------------------------------------------------------------------->
# Connection

The connection to a chTKCat database is achieved by instantiating
*chTKCat* object.

```{r}
k <- chTKCat(
   host="localhost", # default parameter
   port=9101L,       # default parameter
   user="default",   # default parameter
   password=""       # if not provided the password is requested interactively 
)
```

Connection parameters should be adapted according to the ClickHouse database
setup which is documented in the web exploration
interface (The "System" button).

<!----------------------------------------------------------------------------->
<!----------------------------------------------------------------------------->
# Usage

The *chTKCat* object can be used as
a [*TKCat* object](TKCat-User-guide.html#local-tkcat).

```{r, eval=FALSE}
list_MDBs(k)                     # list all the MDBs in a TKCat object
khpo <- get_MDB(k, "HPO")        # get a specific MDBs from the catalog
search_MDB_tables(k, "disease")  # Search table about "disease"
search_MDB_fields(k, "disease")  # Search a field about "disease"
collection_members(k)            # Get collection members of the different MDBs
```

The *khpo* object in the example above is a *chMDB* object.
The data of this type of MDB are located in the ClickHouse database until
the user requests them. They can be use like any other *MDB* object:

- [Getting information](TKCat-User-guide.html#getting-information)
- [Pulling, subsetting and combining](TKCat-User-guide.html#pulling-subsetting-and-combining)
- [Filtering and joining](TKCat-User-guide.html#filtering-and-joining)
- [Merging using collections](TKCat-User-guide.html#merging-with-collections)

The user is able to access only *chMDB* allowed according to his
rights (see the [operation manual][opman]).


<!----------------------------------------------------------------------------->
<!----------------------------------------------------------------------------->
# Reconnecting

If the connection to the database is lost the user can reactivate it by
using the `db_reconnect()` function which works with *chTKCat* objects
but also with *chMDB* objects.


<!----------------------------------------------------------------------------->
<!----------------------------------------------------------------------------->
# Pushing an MDB in a chTKCat instance

MDB can be add to the ClickHouse database of a *chTKCat* object
with the `as_chMDB()` function.

```{r, eval=FALSE}
lk <- TKCat(read_fileMDB(
   path=system.file("examples/ClinVar", package="TKCat")
))
names(lk)
as_chMDB(lk$ClinVar, k, overwrite=FALSE)
```

This function works only if (see the [operation manual][opman]):

- the *ClinVar* database has been initiated by the administrator of the
ClickHouse database,
- the user is allowed to write in the database.




