## Connecting to TKCat after Azure AD authentication in R


### 1. Authentication

- a) Prepare the credentials

```{r}
device_credentials <- AuthHelpers::create_azure_device_credentials(
   resource=c("api://kmt-prd01/user_impersonation"),
   tenant="237582ad-3eab-4d44-8688-06ca9f2e613b",
   app="f55d2b52-9fed-4b05-8b0a-b24cf8149922"
)
```


- b) Print the user code, paste it, go to the verification URL and follow instruction

```{r}
device_credentials$get_user_code()
device_credentials$browse_verification_uri()
```

- c) Request the token

```{r}
device_credentials$request_token()
```

- d) Verify the validity of the token

```{r}
device_credentials$is_valid()
device_credentials$get_token()
device_credentials$get_access_token()
```

### 2. (Optional) Save and reuse credentials

- a) Save the credentials and change permission to limit the access to the file

```{r}
cred_file <- "~/etc/kmt.rds"
saveRDS(device_credentials, file=cred_file, compress=FALSE)
Sys.chmod(cred_file, "0600")
```

- b) In a new session, load the credentials and refresh the token

```{r}
device_credentials <- readRDS("~/etc/kmt.rds")
device_credentials$refresh_token()
```

### 3. Connecting to TKCat with Azure AD token

```{r}
library(TKCat)
k <- chTKCat(
   "tkcat.ucb.com",
   password="",
   port=443, https=TRUE,
   extended_headers=list(
      "Authorization"=paste("Bearer", device_credentials$get_access_token())
   )
)
list_MDBs(k)
```
