## Connecting to knowledge management tools after Azure AD authentication in R

### 0. Requirements

#### TKCat

- From CRAN

```{r, eval=FALSE}
install.packages(TKCat)
```

- From Azure DevOps

```{r, eval=FALSE}
devtools::install_git(
   "git@ssh.dev.azure.com:v3/ucbalm/Translational%20Medicine/TKCat"
)
```


#### AuthHelpers

- From Azure DevOps
<a href="https://dev.azure.com/ucbalm/Translational%20Medicine/_git/AuthHelpers" target="_blank">https://dev.azure.com/ucbalm/Translational%20Medicine/_git/AuthHelpers</a>

```{r, eval=FALSE}
devtools::install_git(
   "git@ssh.dev.azure.com:v3/ucbalm/Translational%20Medicine/AuthHelpers"
)
```

### 1. Authentication

Further details can be found here:
<a href="https://dev.azure.com/ucbalm/Translational%20Medicine/_git/AuthHelpers?path=%2Fsupp%2FUCB-KMT" target="_blank">https://dev.azure.com/ucbalm/Translational%20Medicine/_git/AuthHelpers?path=%2Fsupp%2FUCB-KMT</a>



- a) Generate the credentials from your UCB laptop
(every 3 months after password change)

```{r, eval=FALSE}
library(AuthHelpers)
auth_credentials <- create_azure_authorization_credentials(
   resource = "api://kmt-prd01/user_impersonation",
   tenant = "237582ad-3eab-4d44-8688-06ca9f2e613b",
   app = "f55d2b52-9fed-4b05-8b0a-b24cf8149922"
)
```

- b) Save the credentials and change permission to limit the access to the file

```{r, eval=FALSE}
## Save the credentials
saveRDS(auth_credentials, file="~/etc/kmt_authorization.rds", compress=FALSE)
Sys.chmod("~/etc/kmt_authorization.rds", "0600")
```

You can copy this credentials file in environments from which you want to
access to knowledge management tools (KMT)

### 2. Connecting to KMT with Azure AD token

- a) Load and prepare the credentials

```{r, eval=FALSE}
.get_tk_headers <- do.call(function(){
   credentials <- readRDS("~/etc/kmt_authorization.rds")
   credentials$refresh_token()
   return(function(){
      credentials$refresh_token()
      list(
         "Authorization"=paste("Bearer", credentials$get_access_token())
      )
   })
}, list())
```

- b) Connect to services

```{r, eval=FALSE}
## TKCat
library(TKCat)
.tkcon <- chTKCat(
   "tkcat.ucb.com",
   password="",
   port=443, https=TRUE,
   extended_headers=.get_tk_headers()
)
.db_reconnect <- function(x){
   xn <- deparse(substitute(x))
   nv <- db_reconnect(x, extended_headers=.get_tk_headers())
   assign(xn, nv, envir=parent.frame(n=1))
   invisible(nv)
}
```

```{r, eval=FALSE}
## BED
library(BED)
.reconnectToBed <- function(){
   connectToBed(
      "https://bed.ucb.com",
      remember=FALSE, useCache=TRUE,
      .opts = list(
         extendedHeaders=.get_tk_headers()
      )
   )
}
.reconnectToBed()

## DODO
library(DODO)
.reconnectToDODO <- function(){
   connect_to_dodo(
      "https://dodo.ucb.com",
      remember=FALSE,
      .opts = list(
         extendedHeaders=.get_tk_headers()
      )
   )
}
.reconnectToDODO()
```


Since the Azure AD token expires after one hour, you will have to call
the `.db_reconnect(.tkcon)`, `.reconnectToBed()` and `.reconnectToDODO()`
function after this time.

