
UCB IT Early Solution team made the databases
supporting Knowledge Management Tools (KMT) available
through [Azure Active Directory (AAD)](https://docs.microsoft.com/azure/active-directory/develop/)
authentication. These databases include:

- tkcat.ucb.com for [TKCat](https://patzaw.github.io/TKCat/)
- bed.ucb.com for [BED](https://patzaw.github.io/BED/)
- dodo.ucb.com for [DODO](https://github.com/Elysheba/DODO)

This document describes how to authenticate in R to be able to connect to
these databases from different environments.

## Generate authorization credentials

Authentication within R relies on
the [AzureAuth](https://cran.r-project.org/web/packages/AzureAuth/index.html)
R package and this package must be installed to generate and use credentials.

### From local environment

The *"authorization_code"* method allows to authenticate as a specific user
with relevant access rights. It will open a web browser for authentication
purpose before providing the credentials. Therefore, this method will only
work from local environments and not remote environments such as a linux
workstation.

```{r, eval=FALSE}
library(AzureAuth)
auth_credentials <- AzureAuth::get_azure_token(
   resource = c("api://kmt-prd01/user_impersonation", "offline_access"),
   tenant = "237582ad-3eab-4d44-8688-06ca9f2e613b",
   app = "f55d2b52-9fed-4b05-8b0a-b24cf8149922",
   version=2,
   auth_type="authorization_code",
   use_cache = FALSE
)
```

### From remote environment

For remote environments, you can download the R script generated
automatically with your current credentials using
the "Download credentials R script" button above and source it in R.

To authenticate from a remote environment you can also save the
credentials created locally and copy them in the remote environment.

The credentials created by one or the other method can be refreshed remotely
without having to go through web browser authentication anymore
(see "Connection to KMT section").
**They remain valid about 3 months or until user password changes.**

```{r, eval=FALSE}
saveRDS(auth_credentials, file="~/etc/kmt_authorization.rds", compress=FALSE)
Sys.chmod("~/etc/kmt_authorization.rds", "0600")
```

## Connecting to KMT with saved credentials

The scripts below can be copied in an R script file (e.g. `~/opt/KMT.R`) that
can be sourced whenever needed.

The `.get_tk_headers()` function is made to get an access token from the
saved credentials (see [Credentials section](#credentials)). It is used by
connection calls in the rest of the script but is not supposed to be called
directly.

```{r, eval=FALSE}
## Get updated headers ----
.get_tk_headers <- do.call(function(){
   ## Choose the relevant credentials
   credentials <- readRDS("~/etc/kmt_authorization.rds")
   credentials$refresh()
   return(function(){
      if(!credentials$validate()){
         credentials$refresh()
      }
      list(
         "Authorization"=paste("Bearer", credentials$credentials$access_token)
      )
   })
}, list())
```

The script below shows how to connect to TKCat. It creates 3 objects:

- `.tkcon`: the TKCat connection
- `.db_reconnect()`: a function which should be used instead of
the `TKCat::db_reconnect()` for leveraging credentials
with the `.get_tk_headers()` function described above
- `.tbkm`: a KMR object with TBKM specifications and helper functions
(refer to the relevant documentation)

```{r, eval=FALSE}
## TKCat connection ----
suppressPackageStartupMessages(library(TKCat))
.tkcon <- chTKCat(
   "tkcat.ucb.com",
   password="",
   port=443, https=TRUE,
   extended_headers=.get_tk_headers()
)
.db_reconnect <- function(x){
   xn <- deparse(substitute(x))
   nv <- db_reconnect(x, extended_headers=.get_tk_headers())
   assign(xn, nv, envir=parent.frame(n=1))
   invisible(nv)
}
.tbkm <- get_KMR(.tkcon, "TBKM", check=FALSE)
```

The script below shows how to connect to BED by providing
the `.reconnectToBed()` function.

```{r, eval=FALSE}
## BED connection ----
suppressPackageStartupMessages(library(BED))
.reconnectToBed <- function(){
   connectToBed(
      "https://bed.ucb.com",
      remember=FALSE, useCache=TRUE,
      .opts = list(
         extendedHeaders=.get_tk_headers()
      )
   )
}
.reconnectToBed()
```

The script below shows how to connect to DODO by providing
the `.reconnectToDodo()` function.

```{r, eval=FALSE}
## DODO connection ----
suppressPackageStartupMessages(library(DODO))
.reconnectToDodo <- function(){
   connect_to_dodo(
      "https://dodo.ucb.com",
      remember=FALSE,
      .opts = list(
         extendedHeaders=.get_tk_headers()
      )
   )
}
.reconnectToDodo()
```
